---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# Package **_bio_**

```{r README-0, include=FALSE}
options(bio.use_local_list = TRUE)

options(
  repos = "https://cran.rstudio.com/",
  pkgType = "both",
  install.packages.check.source = "yes",
  install.packages.compile.from.source = "always"
)
```


<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/bio)](https://CRAN.R-project.org/package=bio)
[![GitHub version](https://img.shields.io/badge/GitHub-v`r packageVersion("bio")`-brightgreen.svg)](https://github.com/mokymai/bio)
[![Travis build status](https://travis-ci.com/mokymai/bio.svg?branch=master)](https://travis-ci.com/mokymai/bio)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/mokymai/bio?branch=master&svg=true)](https://ci.appveyor.com/project/mokymai/bio)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Documented on](https://img.shields.io/badge/Documentation-`r gsub('-', '--', Sys.Date())`-yellowgreen.svg)]()
<!-- badges: end -->


Convenience functions to facilitate installation and management of required resources for __*Biostatistics*__ course unit (*BS-2020*).


<center>
<font color="red" size=6>
<br>
Work is still in progress...
</font>
<br>
<font color="darkred" size=4>
Tested only on <em>Windows 10</em> (64-bit) with RStudio 1.2.5033.
</font>
<br><br>
</center>

Bug reports and your feedback are welcome at https://github.com/mokymai/bio/issues.
In the report, indicate your operating system.


# Install

To install this package from GitHub, use code:

```{r README-1, eval=FALSE}
if (!require(remotes)) {install.packages("remotes")}
remotes::install_github("mokymai/bio")
```

# Examples


## Check information about OS, programs and packages

Check information about operating system (OS) and R-related user information.

```{r README-6, paged.print=FALSE, eval=FALSE}
bio::check_user_info()
## Operating system    Windows 10 x64 (build 18362)
## Platform            x86_64-w64-mingw32/x64 (64-bit)
## USERNAME            User
## USERPROFILE         C:/Users/User
## HOME                C:/Users/User/Documents
## R_USER              C:/Users/User/Documents
## R_HOME              C:/PROGRA~1/R/R-36~1.2
## R_LIBS_USER         C:/R/win-library/3.6
```

Check installed R-related programs. 
Options:

- "main" (default) -- checks "R", "RStudio", "Rtools" ("Windows" only) and "XQuartz" ("Mac" only).
- "all" -- additionally checks for presence of "Atom", "Git" and "GitHub".


```{r README-7, eval=FALSE}
bio::check_installed_programs("all")
## ✔ Program R (3.6.2) is installed (recommended 3.6.2, available 3.6.2)
## ✖ Program RStudio is not installed or is not running.
## ✔ Program Rtools is installed.
## ✔ Program Atom is installed.
## ✔ Program Git is installed.
## ✔ Program Meld is installed.
```


Check for installed packages. 
By default, only missing packages or packages with lower than recommended versions are displayed.
Currently available lists are:

- "mini"
- "addins-rmd"
- "bio"
- "bs-2020-initial"
- "gmc-r209"
- "rcmdr"
- "rcmdr-biostat"
- "snippets"
- "spelling-grammar"

```{r README-8a, eval=FALSE, paged.print=FALSE}
bio::check_installed_packages(list_name = "mini")
## ✔ The required versions of all 19 packages (from list 'mini') are already installed.
```

```{r README-8b, eval=FALSE, paged.print=FALSE}
bio::check_installed_packages(list_name = "mini", include = "newer_on_cran")
## ✔ The required versions of all 19 packages (from list 'mini') are already installed.
## ● 1 package has newer version on CRAN.
```


## Update packages

These are the convenience functions to update certain R packages.
The functions try to restart R if used in RStudio.


Update package **bio**.
```{r README-3, eval=FALSE}
bio::update_pkg_bio()
```

Update package **snippets**.
```{r README-4, eval=FALSE}
bio::update_pkg_snippets()
```

Update package **RcmdrPlugin.biostat**.
```{r README-5, eval=FALSE}
bio::update_pkg_rcmdr_biostat()
```



## RStudio settings and user preferences

> You should use functions that reset key bindings (shortcut keys), 
  RStudio settings and snippets **at your own risk**.

### Dictionaries

Update dictionaries.

To download **hunspell** spelling checking dictionaries (including Lithuanian) for RStudio.
```{r eval=FALSE}
bio::download_spellcheck_dictionaries()
```



### Shortcut keys

Change shortcut keys.
Options:

- "bio-default" -- shortcut keys, recommended for BS-2020.

```{r README-9, eval=FALSE}
bio::set_rstudio_keybindings("bio-default")
```

: **Table 1.** Some of shortcut keys set with option "bio-default".

Shortcut  | Action
--------- | -------------------------------------------------------------------------------------------------------------------------------------------
<kbd>Ctrl</kbd>+<kbd>Shift</kbd>+<kbd>Alt</kbd>+<kbd>S</kbd>  | Check spelling in selected text (if package [**wellspell.addin**](https://github.com/nevrome/wellspell.addin){target="_blank"} is installed).
<kbd>Ctrl</kbd>+<kbd>Shift</kbd>+<kbd>Alt</kbd>+<kbd>A</kbd>  | Use "Tidyverse" style on selected lines of R code (*R code*).
<kbd>Ctrl</kbd>+<kbd>Shift</kbd>+<kbd>D</kbd>                 | Duplicate selected text.
<kbd>Ctrl</kbd>+<kbd>Shift</kbd>+<kbd>Alt</kbd>+<kbd>P</kbd>  | Align selected lines at pattern.
<kbd>Ctrl</kbd>+<kbd>Shift</kbd>+<kbd>;</kbd>                 | Insert lower opening quotes `„`.
<kbd>Ctrl</kbd>+<kbd>Shift</kbd>+<kbd>'</kbd>                 | Insert upper closing quotes `“`.
<kbd>Ctrl</kbd>+<kbd>E</kbd>                                  | **[!]** Insert/Enclose with R code block (*R Markdown*).
<kbd>Ctrl</kbd>++<kbd>Alt</kbd>+<kbd>T</kbd>                  | Insert inline code highlighted as R code (R Markdown).
<kbd>Shift</kbd>+<kbd>Alt</kbd>+<kbd>S</kbd>                  | Insert inline LaTeX equation (*R Markdown*).
<kbd>Ctrl</kbd>+<kbd>Alt</kbd>+<kbd>NumPad0</kbd>             | Remove heading/section name (*R Markdown*).
<kbd>Ctrl</kbd>+<kbd>Alt</kbd>+<kbd>NumPad1</kbd>,<br> <kbd>Ctrl</kbd>+<kbd>Alt</kbd>+<kbd>NumPad2</kbd>,<br> <kbd>Ctrl</kbd>+<kbd>Alt</kbd>+<kbd>NumPad3</kbd>,<br> <kbd>Ctrl</kbd>+<kbd>Alt</kbd>+<kbd>NumPad4</kbd>,<br> <kbd>Ctrl</kbd>+<kbd>Alt</kbd>+<kbd>NumPad5</kbd>             | Replace line into/Insert heading (section name) of level 1, 2, 3, 4, or level 5 (*R Markdown*).
<kbd>Ctrl</kbd>+<kbd>Alt</kbd>+<kbd>NumPad6</kbd>             | Insert operator `%in%` (*R code*).

<!-- "addins.rmd::rmd_list_numbered"                : "Ctrl+Shift+Alt+L", -->
<!-- "addins.rs::rs_align_code_at_arrow"            : "Ctrl+Shift+Alt+[", -->
<!-- "addins.rs::rs_align_code_at_equal"            : "Ctrl+Shift+Alt+]", -->
<!-- "addins.rs::rs_insert_exposition_pipe"         : "Ctrl+Alt+S", -->
<!-- "addins.rs::rs_insert_infix_in"                : "Ctrl+Alt+NumPad6", -->
<!-- "addins.rs::rs_insert_line_ds"                 : "Ctrl+Shift+Alt+Q", -->
<!-- "addins.rs::rs_insert_line_ss"                 : "Shift+Alt+Q", -->
<!-- "addins.rs::rs_insert_line_sw"                 : "Ctrl+Shift+Q", -->
<!-- "addins.rs::rs_insert_matrix_multiplication"   : "Ctrl+Alt+NumPad8", -->
<!-- "addins.rs::rs_insert_tee_pipe"                : "Ctrl+Alt+.", -->
<!-- "addins.rs::rs_insert_update_pipe"             : "Ctrl+Alt+,", -->
<!-- "addins.rs::rs_replace_slash_b2fw"             : "Ctrl+Alt+/", -->
<!-- "addins.rs::rs_replace_slash_bd2s"             : "Ctrl+Shift+Alt+\\", -->
<!-- "addins.rs::rs_replace_slash_bs2d"             : "Ctrl+Alt+\\", -->
<!-- "addins.rs::rs_replace_slash_fw2b"             : "Ctrl+Shift+Alt+/", -->
<!-- "bookdown::mathquill"                          : "Ctrl+Alt+NumPad9", -->

More on [Customizing Keyboard Shortcuts](https://support.rstudio.com/hc/en-us/articles/206382178-Customizing-Keyboard-Shortcuts){target="_blank"}.


### RStudio settings 

Reset RStudio user preferences and settings.
Options:

- "bio-default" -- options recommended for course BS-2020.
- "rstudio-default" -- default RStudio settings.

```{r README-10, eval=FALSE}
bio::reset_rs_user_settings(to = "bio-default")
```


### Snippets

Replace R code and R Markdown snippets with those in package "snippets".
More information on 
[Code Snippets](https://support.rstudio.com/hc/en-us/articles/204463668-Code-Snippets){target="_blank"} and on package
[„snippets“](https://gegznav.github.io/snippets/){target="_blank"}.


```{r, eval=FALSE}
snippets::install_snippets_from_package("snippets", type = c("r", "markdown"), backup = TRUE)
```


### Restart R and RStudio

Reload RStudio to see the effect of changed RStudio settings.
(In some cases it is recommended to close and reopen RStudio).

```{r, eval=FALSE}
bio::reload_rstudio()
```

Restart R inside RStudio session.
Useful before installation of R packages.
```{r, eval=FALSE}
bio::restart_r()
```


## RStudio projects

Open (recently used) RStudio project.
```{r README-11, eval=FALSE}
bio::open_project()
## 
## Choose the name of the project (press 0 to cancel): 
## 
##  1: project-1 
##  2: _learn
##  3: biostatistics
##  

##  Selection: 0
##  Cancelled by user.
```

Copy the recent projects to your personal project list.
```{r README-12, eval=FALSE}
bio::update_personal_proj_list()
```
