---
title: "Package **_bio_**"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- # Package **_bio_** -->

```{r README-0, include=FALSE}
options(bio.use_local_list = TRUE)

options(
  repos = "https://cran.rstudio.com/",
  pkgType = "both",
  install.packages.check.source = "yes",
  install.packages.compile.from.source = "always"
)

Sys.setenv(R_REMOTES_NO_ERRORS_FROM_WARNINGS = "true")
```


<!-- badges: start -->
[![GitHub last commit](https://img.shields.io/github/last-commit/mokymai/bio)](https://github.com/mokymai/bio)
[![R build status](https://github.com/mokymai/bio/workflows/R-CMD-check/badge.svg)](https://github.com/mokymai/bio/actions)
[![Codecov test coverage](https://codecov.io/gh/mokymai/bio/branch/master/graph/badge.svg)](https://codecov.io/gh/mokymai/bio?branch=master)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

<!-- [![Documented on](https://img.shields.io/badge/Documentation-`r gsub('-', '--', Sys.Date())`-yellowgreen.svg)]() -->


Convenience functions to facilitate installation and management of required resources for __*Biostatistics*__ course unit (*BS-2020*).


<center>
<!-- <font color="red" size=6> -->
<!-- <br> -->
<!-- Work is still in progress... -->
<!-- </font> -->
<br>
<font color="darkred" size=4>
This package is tested on <em>Windows 10</em> (64-bit) with RStudio 1.2.5033.
</font>
<br><br>
</center>

Bug reports and your feedback are welcome at https://github.com/mokymai/bio/issues.
In the report, indicate your operating system.


# Install

To install this package from GitHub, use code:

```{r README-1, eval=FALSE}
if (!require(remotes)) {install.packages("remotes")}
remotes::install_github("mokymai/bio", dependencies = TRUE)
```

# Examples

## Check information about OS and user

Check information about operating system (OS) and R-related user information.

```{r README-6, paged.print=FALSE, eval=FALSE}
bio::check_user_info()
## Operating system    Windows 10 x64 (build 18362)
## Platform            x86_64-w64-mingw32/x64 (64-bit)
## LOGNAME
## USERNAME            User
## USERPROFILE         C:/Users/User
## HOME                C:/Users/User/Documents
## R_USER              C:/Users/User/Documents
## R_HOME              C:/PROGRA~1/R/R-36~1.2
## R_LIBS_USER         C:/R/win-library/3.6
```

## Check information about programs

Check versions of presence of installed R-related and other programs. 
Options:

- `"main"`{.r} (default) -- checks for "R", "RStudio", "Rtools" ("Windows" only) or "Compiler" (non-"Windows" only) and "XQuartz" ("Mac" only).
- `"all"`{.r} -- additionally checks for presence of "Atom", "Git" and (sometimes) "Meld". 
  <!-- This check may not work properly on "Mac" and "Linux". -->


```{r README-7a, eval=FALSE}
bio::check_installed_programs()
## ✔ Program R (3.6.3) is installed (recommended 3.6.3, available 3.6.3)
## ✖ Program RStudio is not installed or is not running.
## ✔ Program Rtools is installed.
```


```{r README-7b, eval=FALSE}
bio::check_installed_programs("all")
## ✔ Program R (3.6.3) is installed (recommended 3.6.3, available 3.6.3)
## ✔ Program RStudio (1.2.5033) is installed (recommended 1.2.5033, available 1.2.5042)
## ✔ Program Rtools is installed.
## ✖ Program Atom is not detected or configured incorrectly.
## ✔ Program Git is installed.
## ✔ Program Meld is installed.
```


## Check information about "R" packages

The following functions may be used to check installed packages by topic.
By default, only missing packages or packages with lower versions than recommended are displayed.
Currently available lists can be returned by function `bio::get_pkg_lists_local()`{.r} and they include:

```{r echo=FALSE, results='asis'}
res <- paste0(
  "\n",
  paste0("- `'", bio::get_pkg_lists_local(), "'`", collapse = "\n"),
  "\n\n"
)
cat(res)
```

```{r README-8a, eval=FALSE, paged.print=FALSE}
bio::check_installed_packages(list_name = "mini")
## ✔ Minimal required versions of all 19 packages (from list 'mini') are already installed.
```

```{r README-8b, eval=FALSE, paged.print=FALSE}
bio::check_installed_packages(list_name = "mini", include = "newer_on_cran")
## ✔ Minimal required versions of all 19 packages (from list 'mini') are already installed.
## i Note: 1 package has newer version on CRAN.
```


## Update packages

These are the convenience functions to update certain R packages.
The functions restart R before installation if used in RStudio.


Update package **bio**.
```{r README-3a, eval=FALSE}
bio::update_pkg_bio()
```

Check if there are updates for **bio** or its dependency packages.
```{r README-3b, eval=FALSE}
bio::check_updates_bio()
```
```
✔ Minimal required versions of all 24 packages (from list 'bio') are already installed.
```

Update package **snippets**.
```{r README-4, eval=FALSE}
bio::update_pkg_snippets()
```

Update package **RcmdrPlugin.biostat**.
```{r README-5a, eval=FALSE}
bio::update_pkg_rcmdr_biostat()
```


## RStudio settings and user preferences

> Make sure that you understand what you do when you use the functions that 
  reset RStudio key bindings (shortcut keys), settings or snippets. 
> **You may loose all your previous settings.**


### Dictionaries

This function downloads or updates **hunspell** spelling checking dictionaries dictionaries (including Lithuanian) that can be used by RStudio or by tools from **wellspell.addin** package.

```{r eval=FALSE}
bio::download_spellcheck_dictionaries()
```



### Shortcut keys

This function changes shortcut keys in RStudio.

Run function without arguments to see the available options.
```{r README-9a, eval=FALSE, error=TRUE}
bio::reset_rstudio_keybindings()
```
```
Error: The set of RStudio shortcut keys is not defined (argument 'to').
Possible options: 'bio-default', 'rstudio-default'.
```


Options:

- `"bio-default"`{.r} -- shortcut keys recommended for course BS-2020.
- `"rstudio-default"`{.r} -- RStudio defaults.

```{r README-9b, eval=FALSE}
bio::reset_rstudio_keybindings(to = "bio-default")
```

: **Table 1.** Some examples of shortcut keys set with option `"bio-default"`{.r}.

Shortcut  | Action
--------- | -------------------------------------------------------------------------------------------------------------------------------------------
<kbd>Ctrl</kbd>+<kbd>Shift</kbd>+<kbd>Alt</kbd>+<kbd>S</kbd>  | Check spelling in selected text. Package [**wellspell.addin**](https://github.com/nevrome/wellspell.addin){target="_blank"} must be installed.
<kbd>Shift</kbd>+<kbd>Alt</kbd>+<kbd>A</kbd>                  | Use "Tidyverse" style on selected lines of R code. Package **styler** must be installed (*R code*).
<kbd>Ctrl</kbd>+<kbd>Shift</kbd>+<kbd>D</kbd>                 | Duplicate selected text.
<kbd>Ctrl</kbd>+<kbd>Shift</kbd>+<kbd>Alt</kbd>+<kbd>P</kbd>  | Align selected lines at pattern.
<kbd>Ctrl</kbd>+<kbd>Shift</kbd>+<kbd>;</kbd>                 | Insert lower opening quotes `„`.
<kbd>Ctrl</kbd>+<kbd>Shift</kbd>+<kbd>'</kbd>                 | Insert upper closing quotes `“`.
<kbd>Ctrl</kbd>+<kbd>E</kbd>                                  | Insert/Enclose with R code block (*R Markdown*) **[!]** .
<kbd>Ctrl</kbd>+<kbd>Alt</kbd>+<kbd>T</kbd>                   | Insert inline code highlighted as R code (R Markdown).
<kbd>Shift</kbd>+<kbd>Alt</kbd>+<kbd>S</kbd>                  | Insert inline LaTeX equation (*R Markdown*).
<kbd>Ctrl</kbd>+<kbd>Alt</kbd>+<kbd>NumPad0</kbd>             | Remove heading/section name (*R Markdown*).
<kbd>Ctrl</kbd>+<kbd>Alt</kbd>+<kbd>NumPad1</kbd>             | Replace line into/Insert heading (section name) of level 1 (*R Markdown*).
<kbd>Ctrl</kbd>+<kbd>Alt</kbd>+<kbd>NumPad2</kbd>             | ... of level 2 (*R Markdown*).
<kbd>Ctrl</kbd>+<kbd>Alt</kbd>+<kbd>NumPad3</kbd>             | ... of level 3 (*R Markdown*).
<kbd>Ctrl</kbd>+<kbd>Alt</kbd>+<kbd>NumPad4</kbd>             | ... of level 4 (*R Markdown*).
<kbd>Ctrl</kbd>+<kbd>Alt</kbd>+<kbd>NumPad5</kbd>             | ... of level 5 (*R Markdown*).
<kbd>Ctrl</kbd>+<kbd>Alt</kbd>+<kbd>NumPad6</kbd>             | Insert operator `%in%` (*R code*).

<!-- "addins.rmd::rmd_list_numbered"                : "Ctrl+Shift+Alt+L", -->
<!-- "addins.rs::rs_align_code_at_arrow"            : "Ctrl+Shift+Alt+[", -->
<!-- "addins.rs::rs_align_code_at_equal"            : "Ctrl+Shift+Alt+]", -->
<!-- "addins.rs::rs_insert_exposition_pipe"         : "Ctrl+Alt+S", -->
<!-- "addins.rs::rs_insert_infix_in"                : "Ctrl+Alt+NumPad6", -->
<!-- "addins.rs::rs_insert_line_ds"                 : "Ctrl+Shift+Alt+Q", -->
<!-- "addins.rs::rs_insert_line_ss"                 : "Shift+Alt+Q", -->
<!-- "addins.rs::rs_insert_line_sw"                 : "Ctrl+Shift+Q", -->
<!-- "addins.rs::rs_insert_matrix_multiplication"   : "Ctrl+Alt+NumPad8", -->
<!-- "addins.rs::rs_insert_tee_pipe"                : "Ctrl+Alt+.", -->
<!-- "addins.rs::rs_insert_update_pipe"             : "Ctrl+Alt+,", -->
<!-- "addins.rs::rs_replace_slash_b2fw"             : "Ctrl+Alt+/", -->
<!-- "addins.rs::rs_replace_slash_bd2s"             : "Ctrl+Shift+Alt+\\", -->
<!-- "addins.rs::rs_replace_slash_bs2d"             : "Ctrl+Alt+\\", -->
<!-- "addins.rs::rs_replace_slash_fw2b"             : "Ctrl+Shift+Alt+/", -->
<!-- "bookdown::mathquill"                          : "Ctrl+Alt+NumPad9", -->

More on [Customizing Keyboard Shortcuts](https://support.rstudio.com/hc/en-us/articles/206382178-Customizing-Keyboard-Shortcuts){target="_blank"} in RStudio.


### RStudio settings 

This function resets RStudio user preferences and settings.

Run function without arguments to see the available options.
```{r README-10a, eval=FALSE, error=TRUE}
bio::reset_rstudio_user_settings()
```
```
Error: The set of RStudio user settings is not defined (argument 'to').
Possible choices: 'bio-default', 'rstudio-default'.
```


The options:

- `"bio-default"`{.r} -- settings recommended for course BS-2020.
- `"rstudio-default"`{.r} -- default RStudio settings.

```{r README-10b, eval=FALSE}
bio::reset_rstudio_user_settings(to = "bio-default")
```


### Snippets

The function replaces current R code and R Markdown snippets with those in package "snippets".

```{r, eval=FALSE}
snippets::install_snippets_from_package("snippets", type = c("r", "markdown"), backup = TRUE)
```

More information on 
[Code Snippets](https://support.rstudio.com/hc/en-us/articles/204463668-Code-Snippets){target="_blank"} in RStudio and on package
[**snippets**](https://gegznav.github.io/snippets/){target="_blank"}.


### Restart R and RStudio

If you want to see the effect of changed settings, you should reload RStudio.
You may use this function, which reloads RStudio without closing it:

```{r, eval=FALSE}
bio::reload_rstudio()
```
However, in some cases it is recommended to close and reopen RStudio.

Next, to restart R session without closing RStudio, use: 
```{r, eval=FALSE}
bio::restart_r()
```
Useful to apply before trying to install R packages.


<!-- ## RStudio projects -->

<!-- Open (recently used) RStudio project. -->
<!-- ```{r README-11, eval=FALSE} -->
<!-- bio::open_project() -->
<!-- ##  -->
<!-- ## Choose the name of the project (press 0 to cancel):  -->
<!-- ##  -->
<!-- ##  1: project-1  -->
<!-- ##  2: _learn -->
<!-- ##  3: biostatistics -->
<!-- ##   -->

<!-- ##  Selection: 0 -->
<!-- ##  Cancelled by user. -->
<!-- ``` -->

<!-- Copy the recent projects to your personal project list. -->
<!-- ```{r README-12, eval=FALSE} -->
<!-- bio::update_personal_proj_list() -->
<!-- ``` -->
