---
title: "Get started with bio"
vignette: >
  %\VignetteIndexEntry{Get started with bio}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
    eval: false
format:
  html:
    toc: true
---

[![Lifecycle: experimental](lifecycle-experimental.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)

::: {.callout-warning appearance="simple"}
## Lifecycle: experimental

`bio` is **experimental**. It is developed for a specific Biostatistics course
and its user interface may change without a deprecation cycle. Several
functions rewrite files that RStudio owns, so read
[Managing RStudio settings](rstudio-settings.html) before running them on a
machine you care about.
:::

## What `bio` is for

`bio` prepares and inspects the R and RStudio environment used in a
Biostatistics course. It groups into three areas:

| Area | Typical question | Vignette |
|------|------------------|----------|
| Environment checks | *Is my software new enough?* | [Checking your setup](checking-your-setup.html) |
| RStudio configuration | *Can I get the classroom defaults?* | [Managing RStudio settings](rstudio-settings.html) |
| Projects and paths | *Where does RStudio keep this file?* | this vignette |

Nothing in `bio` is required to *use* R — it exists to make a room full of
laptops behave the same way.

## Installation

`bio` and two of its dependencies are published through a
[drat](https://CRAN.R-project.org/package=drat) repository rather than CRAN.
The repository itself is available at
[mokymai/download](https://mokymai.github.io/download/):

```{r}
#| label: install
repos <- c(
  mokymai = "https://mokymai.github.io/download/",
  CRAN = "https://cran.rstudio.com"
)

install.packages("bio", repos = repos)
```

`DESCRIPTION` declares that repository in `Additional_repositories`, so
`install.packages()` resolves `backup.tools` and `snippets` from it too.
See the R manual for details about
[installing packages](https://cran.r-project.org/doc/manuals/r-release/R-admin.html#Installing-packages).

## A short tour

Start by confirming that the surrounding software is present:

```{r}
#| label: tour-check
bio::check_installed_programs()
```

Then look at where RStudio keeps the files `bio` may rewrite:

```{r}
#| label: tour-paths
bio::get_path_rstudio_config_dir()
bio::get_path_rstudio_config_file(which = "current")
bio::get_path_rstudio_keybindings_dir()
bio::get_path_rstudio_snippets_dir()
```

The `open_*()` counterparts open the same locations in a file manager:

```{r}
#| label: tour-open
bio::open_rstudio_config_dir()
bio::open_rstudio_keybindings_dir()
bio::open_backup_dir()
```

## Projects and project lists

RStudio keeps a list of recently opened projects. `bio` can read it, refresh
the user-editable copy, and open a project by a name pattern:

```{r}
#| label: projects
bio::get_path_recent_proj_list()
bio::get_path_user_proj_list()

bio::update_rstudio_proj_list_user()
bio::read_projects(file = bio::get_path_user_proj_list())

bio::open_project("biostat", new_session = TRUE)
```

`open_project()` matches `pattern` against the project names. When several
projects match, it asks which one to open, so it is interactive by design.

## Restarting

```{r}
#| label: restart
bio::restart_r()
bio::restart_rstudio()
bio::rstudio_reload_ui()
```

These require a running RStudio session; outside RStudio they report that and
return without doing anything.

## Where to go next

* [Checking your setup](checking-your-setup.html) — installed versions,
  packages, and what happens when the machine is offline.
* [Managing RStudio settings](rstudio-settings.html) — presets, keybindings,
  dictionaries, backups, and the live-session caveats.

## Related packages and tools

These projects overlap with or complement narrower parts of `bio`:

* [**rstudio.prefs**](https://CRAN.R-project.org/package=rstudio.prefs), by
  S.A. van der Wulp and Daniel D. Sjoberg, is the closest CRAN alternative for
  RStudio preference files and addin shortcuts.
* [**rstudioapi**](https://CRAN.R-project.org/package=rstudioapi), by Kevin
  Ushey, JJ Allaire, Hadley Wickham, and Gary Ritchie, provides the supported
  API for live RStudio integration.
* [**renv**](https://CRAN.R-project.org/package=renv), by Kevin Ushey and
  Hadley Wickham, manages reproducible project-local libraries and lockfiles;
  it complements `bio::list_pkgs_used_in_dir()`.
* [**styler**](https://CRAN.R-project.org/package=styler), by Kirill Müller,
  Lorenz Walthert, and Indrajeet Patil, formats R code.
* [**drat**](https://CRAN.R-project.org/package=drat), by Dirk Eddelbuettel
  and Carl Boettiger, manages CRAN-like package repositories.

For the underlying platforms, use the official
[R manuals](https://cran.r-project.org/manuals.html),
[RStudio IDE User Guide](https://docs.posit.co/ide/user/), and
[Quarto guide](https://quarto.org/docs/guide/).
