---
title: "Checking your setup"
vignette: >
  %\VignetteIndexEntry{Checking your setup}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
    eval: false
format:
  html:
    toc: true
---

[![Lifecycle: experimental](lifecycle-experimental.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)

::: {.callout-warning appearance="simple"}
## Lifecycle: experimental

These helpers are **experimental**. Output formatting and the set of programs
that are checked may change between releases.
:::

## Installed programs

`check_installed_programs()` reports the installed version of R, RStudio,
Quarto, Rtools, and related tools, and — when the machine is online — the
newest version available:

Official installation and release information is available from the
[R Project](https://www.r-project.org/),
[RStudio IDE User Guide](https://docs.posit.co/ide/user/),
[Quarto installation guide](https://quarto.org/docs/get-started/), and
[Rtools documentation](https://cran.r-project.org/bin/windows/Rtools/).

```{r}
#| label: programs
bio::check_installed_programs()
```

`type` selects which group of programs to report on, and
`skip_online_check = TRUE` makes the call purely local:

```{r}
#| label: programs-offline
bio::check_installed_programs(type = "main", skip_online_check = TRUE)
```

### Behaviour without a network

Looking up "the newest available version" needs three things to work: a
reachable network, a reachable endpoint, and a response in the expected shape.
`bio` treats a failure in any of them as *"the available version is unknown"*
rather than as an error:

* no connectivity → a message that the computer is offline;
* the endpoint refuses or times out → a warning naming the tool;
* the page or API answers in an unexpected format → a warning naming the tool.

In every case the installed versions are still reported. A broken or blocked
download never prevents you from seeing what you actually have installed.

### RStudio Desktop specifics

RStudio can be installed for one user or for all users, and `bio` looks in
both kinds of location:

```{r}
#| label: rstudio-scope
install_dir <- bio:::find_rstudio_install_dir()
bio::get_rstudio_install_scope(install_dir)
```

The scope only describes where the *application* lives. RStudio's preferences,
keybindings, and internal state are always per operating-system user.
Posit's guide documents the supported
[RStudio configuration directories](https://docs.posit.co/ide/user/ide/guide/productivity/custom-settings.html).

## Your session

`check_user_info()` collects the details worth pasting into a bug report:

```{r}
#| label: user-info
bio::check_user_info()
```

Smaller helpers answer one question each:

```{r}
#| label: os
bio::get_os_type()
bio::is_64bit_os()
bio::is_32bit_os()
```

## Packages

Check whether specific packages are installed, list everything that is, or
find out which packages a directory of scripts actually uses:

```{r}
#| label: packages
bio::is_pkg_installed(c("dplyr", "ggplot2", "notapackage"))

bio::get_pkgs_installed()

bio::list_pkgs_used_in_dir(path = ".")
bio::list_pkgs_used_in_files("analysis.R")
```

`list_pkgs_used_in_dir()` scans `.R`, `.Rmd`, and `.qmd` files, so it is a
quick way to build the install list for a course folder.
For reproducible project libraries rather than discovery alone, see
[**renv**](https://rstudio.github.io/renv/).

## Versions

Compare two versions, or list the archived versions of a CRAN package:

```{r}
#| label: versions
bio::compare_version(v_installed = "4.5.0", v_required = "4.6.0")

bio::pkg_list_archived_versions("ggplot2")
```

`compare_version()` returns a comparison of two version strings; it is the
helper behind the "is available online" annotations in
`check_installed_programs()`.
