---
title: "Managing RStudio settings"
vignette: >
  %\VignetteIndexEntry{Managing RStudio settings}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
    eval: false
format:
  html:
    toc: true
---

[![Lifecycle: experimental](lifecycle-experimental.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)

::: {.callout-warning appearance="simple"}
## Lifecycle: experimental

The functions on this page **overwrite files that belong to you**: RStudio
preferences, keybindings, and spellcheck dictionaries. They are experimental
and their preset contents change between releases. Back-ups are made by
default — keep it that way until you are sure.
:::

## Preferences

Four presets are available:

| `to` | What it does |
|------|--------------|
| `"rstudio-default"` | RStudio's own defaults |
| `"bio-default"` | course defaults, light theme |
| `"bio-dark-blue"` | course defaults, Cobalt theme |
| `"bio-black"` | course defaults, Chaos theme |

```{r}
#| label: prefs-reset
bio::rstudio_reset_user_settings(to = "bio-default")
```

`backup = TRUE` (the default) copies the current `rstudio-prefs.json` into the
backup directory first, and `ask = TRUE` asks for confirmation. An automated
run needs both switched off:

```{r}
#| label: prefs-reset-auto
bio::rstudio_reset_user_settings(to = "bio-default", backup = TRUE, ask = FALSE)
```

### What happens inside RStudio versus outside it

The two situations genuinely differ, and it is worth knowing which one you are
in:

* **Outside RStudio** (`Rscript`, a terminal, CI) the presets are merged
  straight into `rstudio-prefs.json`. The write is transactional: if any preset
  fails, the original file is restored byte-for-byte, or removed again if there
  was no original.
* **Inside a running RStudio session** the presets go through
  `rstudioapi::writeRStudioPreference()`. RStudio owns the file and re-persists
  its in-memory state, so `bio` does not delete or roll back the file — doing
  so would have no effect anyway.

A preference key that your RStudio version does not recognise is reported in a
single warning and skipped; the remaining keys are still applied. Only an
unreadable preset file aborts the operation.

::: {.callout-note appearance="simple"}
`ask = TRUE` additionally issues RStudio's `clearUserPrefs` command. That
command opens its own confirmation dialog which cannot be suppressed, so it is
deliberately limited to the interactive path.
:::

### Comparing before you change anything

```{r}
#| label: prefs-compare
bio::rstudio_compare_user_settings(to = "bio-default")
bio::rstudio_compare_user_settings(to = "rstudio-default", output = "verbose")
```

`source` chooses where the "current" values come from:

* `"auto"` (default) — live values inside RStudio, the file otherwise;
* `"live"` — `rstudioapi`, and therefore a running session;
* `"file"` — always `rstudio-prefs.json`.

Because that file only stores values you changed from RStudio's built-in
defaults, the file-based comparison fills in unset keys from your local
RStudio installation's preference schema when it can find it.

## Keybindings

```{r}
#| label: keybindings
bio::rstudio_reset_keybindings(to = "bio-default")
bio::rstudio_reset_keybindings(to = "rstudio-default", backup = TRUE)
```

This is a plain file copy, so it works without a running RStudio session.
Restart RStudio afterwards, or call `bio::rstudio_reload_ui()`.

## Spellcheck dictionaries

```{r}
#| label: dictionaries
bio::rstudio_install_spellcheck_dictionaries()
bio::rstudio_delete_spellcheck_dictionaries()
```

Inside RStudio this delegates to the IDE's own downloader. Outside RStudio it
downloads Posit's dictionary archive itself, retrying interrupted transfers and
falling back to the system `curl` command. Before extracting, the archive is
checked: it must contain the expected dictionary files, and any entry that
would write outside the target directory is refused.

The function invisibly returns `TRUE` or `FALSE`, so you can branch on it:

```{r}
#| label: dictionaries-check
ok <- bio::rstudio_install_spellcheck_dictionaries()
if (!ok) message("Dictionaries were not installed.")
```

::: {.callout-important appearance="simple"}
`secure = FALSE` fetches the archive over an unencrypted connection and
extracts it into your configuration directory. It warns, and you should not
need it.
:::

## Back-ups

Every destructive helper writes a timestamped copy first. To find them:

```{r}
#| label: backups
bio::open_backup_dir()
```

Restoring is a manual file copy — deliberately, so that nothing overwrites
your current state without you asking.

## The files themselves

```{r}
#| label: paths
bio::get_path_rstudio_config_dir()
bio::get_path_rstudio_config_file(which = "current")
bio::get_path_rstudio_config_file(which = "bio-default")
bio::get_path_rstudio_keybindings_dir()
bio::get_path_rstudio_snippets_dir()
bio::get_path_rstudio_internal_state_dir()
bio::get_path_r_environ()
```

Open them instead with `bio::open_rstudio_config_file()`,
`bio::open_rstudio_keybindings_dir()`, `bio::open_rstudio_snippets_dir()`, or
`bio::open_r_environ()`.

The configuration directory (`%APPDATA%/RStudio` on Windows,
`~/.config/rstudio` elsewhere) holds preferences, keybindings, snippets, and
dictionaries. The internal state directory (`%LOCALAPPDATA%/RStudio`,
`~/.local/share/rstudio`) holds session state and is not something `bio`
rewrites.
